The human in the loop: volunteer computing as a socio-technical system

### 2016-02-28 Alife16/paper.tex
Updates state of the art
Reducing it at the same time, but adding some heretofore unmentioned works. Closes #101

%\usepackage[pdftex]{graphicx}
\graphicspath{{../img/}}
\DeclareGraphicsExtensions{.pdf}

\title{The human in the loop: volunteer computing as a socio-technical
system}
\author{Juan-J.~Merelo*$^1$, Paloma de las Cuevas $^1$, Pablo
  García-Sánchez $^1$, Mario Garc\'ia-Valdez$^2$\\
\mbox{}\\
$^1$Dept. of Computer Architecture and Technology and CITIC University of Granada \\
$^2$Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana \\
{\tt jmerelo@ugr.es}, {\tt mario@tectijuana.edu.mx}}

@@

In this paper, our intention is to
present the {\sf NodIO} framework and analyze it from two different
angles: the purely technical angle, including the decisions that went
into its features, and the social angle, to account on how the persons
that voluntarily choose to enter the web page running the system
determine its performance and how it can be modeled and, eventually,
predicted.

The rest of the paper is organized as follows: Next we present the
state of the art (Section \ref{sec:soa}) in web-based distributed
computational systems along with attempts to predict and model its behavior. 
Then, a detailed description of the proposed framework is presented
in Section \ref{sec:description}, followed by the baseline experiments
in Section \ref{sec:experiments}, and experiments using a new client
architecture that uses web workers in Section \ref{sec:w2}.
Finally, conclusions and future lines of work are presented in Section
\ref{sec:conclusion}.

@@

Volunteer computing involves an user running a program voluntarily
and, as such, has been deployed in many different ways from the
beginning of the Internet, starting with the SETI@home framework for
processing extraterrestrial signals \cite{david-seti:home}. However
the dual introduction of JavaScript as an universal language for the
browser and the browser as an ubiquitous web and Internet client has
made this combination the most popular for volunteer computing
frameworks such as the one we are using here and whose first version
was described in \cite{DBLP:conf/gecco/GuervosG15}.

JavaScript can be used either for unwitting
\citep{unwitting-ec} or volunteer 
\citep{langdon:2005:metas,gecco07:workshop:dcor} distributed
evolutionary computation and it has been used ever since by several
authors, including more recent efforts
\citep{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210} 
that even
used the client's GPU \citep{duda2013gpu} or create an object-oriented
framework for evolution in the browser \citep{EvoStar2014:jsEO}. Many other researchers have
used Java \citep{chong:1999:jDGPi} and others have gone away from the
server-based paradigm to embrace peer to peer systems
\citep{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These computing
platforms avoid single points of failure (the server) but, since they
need a certain amount of infrastructure installed to start, the
threshold to join them is much lower.

On the other hand, we are also interested in measuring the performance
of volunteer computing systems, an area in which there have been
relatively few efforts.
There were some initial attempts to avoid the differences in performance
that could be obtained from volunteers  by making
the algorithm adaptive to the kind of resources allotted to it
\citep{milani2004online}, which is actually not such a big problem in
algorithms such as the EA that can easily be
parallelized via population splitting or farming out evaluations to all
the nodes available. Lately, several approaches have focused on the
fault-tolerance of volunteer algorithms
\citep{gonzalez2010characterizing} which it can, of course, be studied in
the more general context of distributed computing 
\citep{nogueras2015studying} or including it in a more general study of the
performance of the EA itself
\citep{DBLP:journals/gpem/LaredoBGVAGF14}. This performance cannot be
measured without understanding the dynamics of this kind of systems. Initial
work was done for peer to peer systems by Stutzbach et
al. \citep{stutzbach2006understanding} and extended to volunteer
computing by Laredo et al. \citep{churn08,laredo2008rcp}.

@@

As far as we know, this paper presents one of the few experiments
that measure the performance of a sociotechnical metacomputer. 
The algorithms used, as well as the methodology 
for gathering resources will be described next, 
together with the results obtained in this initial setup.

@@

Although we think that, in general, the results presented in this
paper are independent of the problem chosen, it is true that even if
it is a difficult problem for evolutionary algorithms, it takes a few
seconds with the right settings to solve. This makes it difficult also
for this kind of asynchronous framework, but it would be interesting
to check the performance for other computationally demanding
tasks, especially ones that cover the middle ground between a problem
finishing in seconds or minutes (Trap) and weeks (Rastrigin F15). 
Besides, the two versions of the framework include as an
algorithmic variant using random population size. We do not think this
has had a big influence on the results and in fact this is not noticed
for F15, which is the second version tested but it would be interesting to
measure exactly what this influence has been. In general there are
many issues with the evolutionary algorithm implementation itself,
including using different, or adaptive, policies for inserting and
sending individuals to the pool as was done in \citep{araujo2008mam},
using different policies for population initialization, and also the
incorporation of high-speed local resources to the pool to check what
would be the real influence of the volunteer pool to the final
performance.

>>>>

\graphicspath{{../img/}}
\DeclareGraphicsExtensions{.pdf}

\title{The human in the loop: volunteer computing as a socio-technical
system}
\author{Juan-J.~Merelo*$^1$, Paloma de las Cuevas $^1$, Pablo
  Garc\'ia-S\'anchez $^1$, Mario Garc\'ia-Valdez$^2$\\
\mbox{}\\
$^1$Dept. of Computer Architecture and Technology and CITIC University of Granada \\
$^2$Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana \\
{\tt jmerelo@ugr.es}, {\tt mario@tectijuana.edu.mx}}

@@

In this paper, our intention is to
present the {\sf NodIO} framework and analyze it from two different
angles: the purely technical angle, including the decisions that went
into its features, and the social angle, to account on how the persons
that voluntarily choose to enter the web page running the system
determine its performance and how it can be modeled and, eventually,
predicted. Next we present the state of the art in web-based
volunteer-based systems along with attempts to predict and model its
behavior.

@@

Volunteer computing involves an user running a program voluntarily
and, as such, has been deployed in many different ways from the
beginning of the Internet, starting with the SETI@home framework for
processing extraterrestrial signals \citep{david-seti:home}. However
the dual introduction of JavaScript as an universal language for the
browser and the browser as an ubiquitous web and Internet client has
made this combination the most popular for volunteer computing
frameworks such as the one we are using here and whose first version
was described in \citep{DBLP:conf/gecco/GuervosG15}.

JavaScript can be used either for unwitting
\citep{unwitting-ec,boldrin2007distributed} or volunteer 
\citep{langdon:2005:metas,gecco07:workshop:dcor} distributed
evolutionary computation and it has been used ever since by several
authors, including more recent efforts
\citep{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210}. Many other researchers have
used Java \citep{chong:1999:jDGPi} and others have gone away from the
server-based paradigm to embrace peer to peer systems
\citep{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These computing
platforms avoid single points of failure (the server) but, since they
need a certain amount of infrastructure installed to start, the
threshold to join them is much lower.

On the other hand, we are also interested in measuring the performance
of volunteer computing systems, an area in which there have been
relatively few efforts.
There were some initial attempts to avoid the differences in performance
that could be obtained from volunteers  by making
the algorithm adaptive to the kind of resources allotted to it
\citep{milani2004online}, which is actually not such a big problem in
algorithms such as the EA that can easily be
parallelized via population splitting or farming out evaluations to all
the nodes available. Lately, several approaches have focused on the
fault-tolerance of volunteer algorithms
\citep{gonzalez2010characterizing} which it can, of course, be studied in
the more general context of distributed computing 
\citep{nogueras2015studying} or including it in a more general study of the
performance of the EA itself
\citep{DBLP:journals/gpem/LaredoBGVAGF14}. This performance cannot be
measured without understanding the dynamics of this kind of systems. Initial
work was done for peer to peer systems by Stutzbach et
al. \citep{stutzbach2006understanding} and extended to volunteer
computing by Laredo et al. \citep{churn08,laredo2008rcp}. A similar
study was performed by Martínez et al. on the Capataz system
\citep{martinez2015capataz}; however in this case the number of
computers used was known in advance and the main focus was on
measuring speed up and how job bundling helped to reduce overhead and
enhance performance.

@@

As far as we know, this paper presents one of the few experiments
that measure the performance of a sociotechnical
metacomputer. Apolónia et al. \citep{apolonia2012enhancing} used the
Facebook protocol to distribute tasks among the {\em walls} of
friends, explicitly using the social network for computing. However,
it stopped short of relating performance to the macro measures of the
users' social networks. 
The algorithms used, as well as the methodology 
for gathering resources will be described next, 
together with the results obtained in this initial setup.

@@

Although we think that, in general, the results presented in this
paper are independent of the problem chosen, it is true that even if
it is a difficult problem for evolutionary algorithms, it takes a few
seconds with the right settings to solve. 
Besides, the two versions of the framework include as an
algorithmic variant using random population size. We do not think this
has had a big influence on the results and in fact this is not noticed
for F15, which is the second version tested but it would be interesting to
measure exactly what this influence has been. In general there are
many issues with the evolutionary algorithm implementation itself,
including using different, or adaptive, policies for inserting and
sending individuals to the pool,
using different policies for population initialization, and also the
incorporation of high-speed local resources to the pool to check what
would be the real influence of the volunteer pool to the final
performance.
